Smart Spreading Factor Assignment for LoRaWANs

### 2019-01-09 paper.tex
grammer

\author{\IEEEauthorblockN{Tugrul Yatagan\IEEEauthorrefmark{1} and Sema F. Oktug\IEEEauthorrefmark{2}}
\IEEEauthorblockA{Department of Computer Engineering,
Istanbul Technical University\\
Istanbul, Turkey\\
Email: \IEEEauthorrefmark{1}yatagan@itu.edu.tr, \IEEEauthorrefmark{2}oktug@itu.edu.tr}}
\maketitle

\begin{abstract}
Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas. Focus of this work is one of the most promising LPWAN technologies: LoRa. LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum (CSS). LoRa modulation trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth. Collisions in LoRaWAN networks are strongly related with spreading factor (SF) assignment of nodes which indeed effects network performance. In this work, a simulation environment to study different SF assignment schemes is implemented. Furthermore, novel SF assignment strategy is proposed which utilizes a machine learning technique for optimization. Finally, promising simulation results for proposed machine learning SF assignment technique is presented.
\end{abstract}

@@

\section{Introduction}
\par In the last few years, number of Internet of Things (IoT) applications increased exponentially \cite{7721743}. Recent developments on LPWAN technologies has great impact on growth of number of IoT applications. LPWAN technologies addresses some of the well-known wireless communication challenges. Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication technologies (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time. Cellular networks can provide long range and high data rate but they are complex and consume too much power. Besides, most of the IoT applications do not require high data rate. Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best \cite{7815384}. LPWAN technologies fill the technology gap between short range and cellular technologies by providing low power and long range communication. LPWAN technologies basically sacrifice data rate to provide low power consumption.

\par There are several emerging LPWAN technologies. LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well-known LPWAN technologies. LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost \cite{7815384}. Both LoRa and Sigfox are known for ultra low power consumption and resilience to interference.  While NB-IoT and LTE-M are promoted for higher data rate. LoRa has open standard MAC protocol called LoRaWAN. LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access. LoRaWAN networks can be deployed as a private network like WiFi. However, Sigfox and NB-IoT are only available with operator contract \cite{7815384}. Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink. Also, Sigfox packet payload size is limited to 12 bytes for uplink and 8 bytes for downlink. However LoRa supports up to 243 bytes payload size and NB-IoT supports up to 1600 bytes payload size. Sigfox maximum data rate is 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.

\par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between receive sensitivity and transmission time on air \cite{7803607}. LoRa supports 6 different SF option. Simultaneous same SF transmissions are prone to collision however different SF transmissions in same channel are orthogonal to each other up to some extent, thus SF assignment is crucial for overall network performance. In this work, a custom made LoRa discrete event simulator is introduced to study different LoRa SF assignment strategies. We also proposed an SF assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using the simulator.

\par The following two sections provides an background information about LoRaWAN and LoRa. Section \ref{Other Related Works} describes other related works. Section \ref{Proposed Technique} describes proposed technique for SF assignment. Simulation environments and results are shown in Section \ref{Simulation Environment} and \ref{Simulation Results}. Finally, Section \ref{Conclusion} concludes the paper.

@@

\subsection{End node}
\par LoRaWAN end node (EN) is a low power embedded device that only communicates between gateways. LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C. Different classes provides LPWAN solutions to different deployments and topologies. Class A end nodes generate uplink transmission at any time and only receives a period of time after uplink transmission. Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission. Receive window is synchronized using beacon packet transmitted by gateways. Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission. This provides Class C end nodes to low latency donwlink communication but it requires more power consumption. In this paper, we focused on Class A end devices since Class A behavior provides the lowest power consumption.

\subsection{Gateway}
\par LoRaWAN gateway (GW) is a device that receive/transmit packets coming from/to end nodes. A typical gateway can receive from multiple channels at the same time. Gateways are typically connected to power grid so power consumption of a gateway is insignificant in most of the deployments.

@@

\par When BW and CR are constant, as the SF increases, the data rate decreases. Increasing the SF makes the signal more resilient to noise thus increases the transmission range. Increasing the SF also increases the transmission duration which increases the power consumption. Therefore it is possible to trade between data rate, range and power consumption by changing SF.

@@

\par In a LoRaWAN network, initially, a node is not aware of how far it is from a GW. However, a node can guess the distance from a gateway by observing received signal power of a downlink transmission. If a downlink transmission received signal power is too high then node can decrease its next transmission SF or transmit power to decrease power consumption. This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper. Also a GW can request from a node to decrease node's SF or transmit power.

\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single GW is illustrated. Different color rings represent achievable range of different SFs from the GW and different color circles represent selected SF of the nodes. Most of the end devices which close to the GW will fall into lowest SF (SF7) area section. End devices which close to the GW will probably select lowest SF all the time. Which results a lot of collisions between same SF transmissions. Number of collisions increases while number of end devices close to the GW increases.

@@

\section{Other Related Works} \label{Other Related Works}
The literature related to the work presented in this paper has started growing recently. LPWAN technologies and especially LoRa  attracted researchers attentions lately. We listed some of these works which studies LoRaWAN and LoRa SF.

@@

\par In \cite{s17061193}, the authors investigated single gateway LoRaWAN network scalability in terms of the number of end nodes using a simulation model based on real measurements. They measure the impact of two concurrent LoRa transmissions on each other by using physical LoRaWAN end devices and a gateway then they created a simulation model from their measurements. Their results show that LoRaWAN has better scalability than pure ALOHA since a LoRa packet may still go through under collision if the the last six symbols of preamble and header of the packet does not collide.

\par In \cite{8267219}, the authors studied imperfect orthogonality between different LoRa SF transmissions. The authors state that a LoRa transmission can be interfered even between different SF transmissions when power of the interfering signal significantly overcomes the reference signal. Their experimental results show that this power difference is around 16 dB and this power difference can be seen when a interferer is close to receiver or multiple interferer can create this power difference cumulatively.

@@

\section{Proposed Technique} \label{Proposed Technique}
\par The collision issue illustrated in Figure \ref{fig:collision} can be solved by forcing some of the close nodes to select higher SF even they can able to communicate with lower SF. This may prevent collisions due to the orthogonality of different SF as shown in Figure \ref{fig:collision_solution_single_gw}. Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}. However, how network should assign which SF to which node becomes an another issue. Increasing a node's SF should be done carefully since higher SF means higher time on air and higher time on air is more prone to collision with other higher SF transmissions. In multiple GW scenarios, this approach may increase the collisions with nodes in other GW's range. Thus, extra care must be taken for nodes in intersection area of GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.

\par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes SFs requires different rules. For this reason, we proposed a machine learning based SF assignment scheme to decrease same SF transmission collisions. This technique learns behavior of the nodes in a network. A NS can keep track of successful uplink transmissions and their SFs. NS can also keep track of some of the collided transmission if header part of the packet does not interfered at location of the GW. However, NS cannot keep track of transmissions with receive power lower than the GW sensitivity. Using these obtained information NS can train a classifier for predicting future transmission result for specific node and specific SF. Using this prediction model NS can assign the lowest possible SF to nodes considering the collision probability.

\par In this work, a decision tree classifier is used to predict transmission results. Gini index criteria is used for measuring quality of decision tree splits. As in most of the machine learning applications, increasing the data set size increases the prediction accuracy. For in this case, data set size is directly proportional to simulation duration. Thus increasing simulation duration, increases the prediction accuracy. For real world cases, NS can keep track of transmission and it can create a classifier daily basis, then GWs can request from nodes to use this new SF.

@@

\section{Simulation Environment} \label{Simulation Environment}
\par A discrete event simulator is implemented in Python to study effects of different SF strategies in LoRaWANs. The LoRaWAN SF simulation tool is publicly available at https://github.com/tugrulyatagan/simlorasf. Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies. Simulator can generate uniformly distributed circular shape network topology with radius (m), number of nodes and number of gateways input parameters. Global simulation input parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and spreading factor assignment method. With these inputs, the simulator provide outputs of total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).

@@

\subsection{Machine Learning Model}
\par We have been able to generate mass amount of LoRaWAN transmission logs for different topologies using the simulator. To come up with a solution for SF assignment issue, a machine learning prediction scheme is integrated to the simulator. A classifier is trained from transmission data and this classifier is used for selecting best possible SF for the nodes. Simulator first runs a simulation with assigning random SF to nodes. Simulation results are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission. Transmission result can be successful, interfered or under sensitivity. These data is fed to Python scikit-learn decision tree classifier. After the classifier model is generated, a second simulation is run with classifier. Simulator selects the lowest possible SF considering  prediction of transmission. If a transmission of a node with lowest possible SF is predicted as interfered transmission by classifier, then simulator assigns a higher SF. If no SF is predicted as successful, then simulator again assigns lowest possible SF.

@@

\par In Figure \ref{fig:sf_pdr}, PDR plot of different SF assignment is shown. Randomly generated network topology radius is set to 3000 meters, number of GW is set to 1 and packet generation rate is set to 0.01 packet per second. Increasing SFs, increases time on air. This increases the number of collisions thus decreases the PDR of the network. PDR of higher SFs are decreased to almost 0 while number of node is increasing. Since network topology radius is quite small almost all SFs can reach to the GW.

@@

\par In Figure \ref{fig:predict_pdr}, PDR plot of machine learning prediction scheme is shown. Randomly generated network topology radius is set to 5000 meters, number of GW is set to 3 and packet generation rate is set to 0.01 packet per second. Prediction model needs nodes location so three GWs is used for prediction SF scheme since three GWs is enough to roughly locate nodes positions by triangulation. Prediction scheme gives better PDR when number of nodes increases since number of interference is also increases. Prediction scheme can only improve network performance when LoRa interference is present. Also prediction scheme gives better results when nodes are deployed to close to the gateway, since nodes have margin to increase their SFs. If a node is far away from the GW, then prediction scheme cannot increase the SF to avoid interference since the SF is already high.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{predict_pdr_r5000_g3_p0.01_s3600}.png}
\caption{PDR of prediction SF assignment scheme.}
\label{fig:predict_pdr}
\end{figure}

@@

\par As for future works, transmit power optimization can be included to proposed machine learning scheme. In this paper, we assume that nodes always use maximum transmit power for uplink transmission, however nodes close to the GW can decrease transmit power to save energy. This will make transmissions to more vulnerable to interference thus requires extra care. Optimizing another node parameter alongside SF can be interesting research. Also other machine learning methods can be investigated for SF optimization. Reinforcement learning is a good candidate for this network optimization issue.

>>>>

\author{\IEEEauthorblockN{Tugrul Yatagan and Sema F. Oktug}
\IEEEauthorblockA{Department of Computer Engineering,
Istanbul Technical University\\
Istanbul, Turkey\\
Email: \{yatagan, oktug\}@itu.edu.tr}}
\maketitle

\begin{abstract}
Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas. Focus of this work is one of the most promising LPWAN technologies: LoRa. LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum (CSS). LoRa modulation trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth. Collisions in LoRaWAN networks are strongly related with spreading factor (SF) assignment of nodes which indeed effects network performance. In this work, a simulation environment to study different SF assignment schemes is implemented. Furthermore, a novel SF assignment strategy is proposed which utilizes a machine learning technique for optimization of SF assignment. Finally, promising simulation results for proposed machine learning SF assignment technique is presented.
\end{abstract}

@@

\section{Introduction}
\par In the last few years, number of Internet of Things (IoT) applications increased exponentially \cite{7721743}. Recent developments on LPWAN technologies has great impact on growth of number of IoT applications. LPWAN technologies address some of the well-known wireless communication challenges. Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication technologies (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time. Cellular networks can provide long range and high data rate, but they are complex and consume too much power. Besides, most of the IoT applications do not require high data rate. Short-range communication methods can provide relatively low power consumption, but their range is limited to a few hundred meters at best \cite{7815384}. LPWAN technologies fill the technology gap between short range and cellular technologies by providing low power and long-range communication. LPWAN technologies basically sacrifice data rate to provide low power consumption.

\par There are several emerging LPWAN technologies. LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well-known LPWAN technologies. LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost \cite{7815384}. Both LoRa and Sigfox are known for ultra-low power consumption and resilience to interference.  While NB-IoT and LTE-M are promoted for higher data rate. LoRa has open standard MAC protocol called LoRaWAN. LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access. LoRaWAN networks can be deployed as a private network like WiFi. However, Sigfox and NB-IoT are only available with operator contract \cite{7815384}. Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink. Also, Sigfox packet payload size is limited to 12 bytes for uplink and 8 bytes for downlink. However, LoRa supports up to 243 bytes payload size and NB-IoT supports up to 1600 bytes payload size. Sigfox maximum data rate is 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.

\par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between receive sensitivity and transmission time on air \cite{7803607}. LoRa supports 6 different SF option. Simultaneous same SF transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other up to some extent, thus SF assignment is crucial for overall network performance. In this work, a custom made LoRa discrete event simulator is introduced to study different LoRa SF assignment strategies. We also proposed an SF assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using the simulator.

\par The following two sections provides a background information about LoRaWAN and LoRa. Section \ref{Other Related Works} describes other related works. Section \ref{Proposed Technique} describes proposed technique for SF assignment. Simulation environments and results are shown in Section \ref{Simulation Environment} and \ref{Simulation Results}. Finally, Section \ref{Conclusion} concludes the paper.

@@

\subsection{End node}
\par LoRaWAN end node (EN) is a low power embedded device that only communicates to gateways. LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C. Different classes provides LPWAN solutions to different deployments and topologies. Class A end nodes generate uplink transmission at any time and only receives a period of time after uplink transmission. Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission. Receive window is synchronized using a beacon packet transmitted by gateways. Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission. This provides Class C end nodes to low latency downlink communication, but it requires more power consumption. In this paper, we focused on Class A end devices since Class A behavior provides the lowest power consumption.

\subsection{Gateway}
\par LoRaWAN gateway (GW) is a device that receive/transmit packets coming from/to end nodes. A typical gateway can receive from multiple channels at the same time. Gateways are typically connected to power grid, so power consumption of a gateway is insignificant in most of the deployments.

@@

\par When BW and CR are constant, as the SF increases, the data rate decreases. Increasing the SF makes the signal more resilient to noise thus increases the transmission range. Increasing the SF also increases the transmission duration which increases the power consumption. Therefore, it is possible to trade between range and power consumption by changing SF.

@@

\par In a LoRaWAN network, initially, a node is not aware of how far it is from a GW. However, a node can guess the distance from a gateway by observing received signal power of a downlink transmission. If a downlink transmission received signal power is too high, then node can decrease its next transmission SF to decrease power consumption. This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper. Also, a GW can request from a node to decrease node's SF or transmit power.

\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single GW is illustrated. Different color rings represent achievable range of different SFs from the GW and different color circles represent selected SF of the nodes. Most of the end devices which close to the GW will fall into lowest SF (SF7) area section. End devices which close to the GW will probably select lowest SF all the time. This results a lot of collisions between same SF transmissions. Number of collisions increases while number of end devices close to the GW increases.

@@

\section{Other Related Works} \label{Other Related Works}
\par The literature related to the work presented in this paper has started growing recently. LPWAN technologies and especially LoRa attracted researchers’ attentions lately. We listed some of these works which studies LoRaWAN and LoRa SF.

@@

\par In \cite{s17061193}, the authors investigated single gateway LoRaWAN network scalability in terms of the number of end nodes using a simulation model based on real measurements. They measure the impact of two concurrent LoRa transmissions on each other by using physical LoRaWAN end devices and a gateway then they created a simulation model from their measurements. Their results show that LoRaWAN has better scalability than pure ALOHA since a LoRa packet may still go through under collision if the last six symbols of preamble and header of the packet does not collide.

\par In \cite{8267219}, the authors studied imperfect orthogonality between different LoRa SF transmissions. The authors state that a LoRa transmission can be interfered even between different SF transmissions when power of the interfering signal significantly overcomes the reference signal. Their experimental results show that this power difference is around 16 dB and this power difference can be seen when an interferer is close to a receiver or sum of multiple interferer energy can create this power difference.

@@

\section{Proposed Technique} \label{Proposed Technique}
\par The collision issue illustrated in Figure \ref{fig:collision} can be solved by forcing some of the close nodes to select higher SF even they can able to communicate with lower SF. This may prevent collisions due to the orthogonality of different SF as shown in Figure \ref{fig:collision_solution_single_gw}. Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}. However, how network should assign which SF to which node becomes another issue. Increasing a node's SF should be done carefully since higher SF means higher time on air and higher time on air is more prone to collision with other higher SF transmissions. In multiple GW scenarios, this approach may increase the collisions with nodes in other GW's range. Thus, extra care should be taken for nodes in intersection area of GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.

\par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes SFs requires different rules. For this reason, we proposed a machine learning based SF assignment scheme to decrease same SF transmission collisions. This technique learns behavior of the nodes in a network. A NS can keep track of successful uplink transmissions and their SFs. NS can also keep track of some of the collided transmission if header part of the packet is not interfered at location of the GW. However, NS cannot keep track of transmissions with receive power lower than the GW sensitivity. Using these obtained information, NS can train a classifier for predicting future transmission result for specific node and specific SF. Using this prediction model NS can assign the lowest possible SF to nodes considering the collision probability.

\par In this work, a decision tree classifier is used to predict transmission results. Gini index criteria is used for measuring quality of decision tree splits. As in most of the machine learning applications, increasing the data set size increases the prediction accuracy. For in this case, data set size is directly proportional to simulation duration. Thus, increasing simulation duration, increases the prediction accuracy. For real world cases, NS can keep track of transmission and it can create a classifier daily basis, then GWs can request from nodes to use this new SF.

@@

\section{Simulation Environment} \label{Simulation Environment}
\par A discrete event simulator is implemented in Python to study effects of different SF strategies in LoRaWANs. The LoRaWAN SF simulation tool source code is available at \cite{simlorasf}. Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies. Simulator can generate uniformly distributed circular shape network topology with radius (m), number of nodes and number of gateways input parameters. Global simulation input parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and spreading factor assignment method. With these inputs, the simulator provides outputs of total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).

@@

\subsection{Machine Learning Model}
\par We have been able to generate mass amount of LoRaWAN transmission logs for different topologies using the simulator. To come up with a solution for SF assignment issue, a machine learning prediction scheme is integrated to the simulator. A classifier is trained from transmission logs and this classifier is used for selecting best possible SF for the nodes. Simulator first runs a simulation with assigning random SF to nodes. Simulation results are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission. Transmission result can be successful, interfered or under sensitivity. These data are fed to Python scikit-learn decision tree classifier. After the classifier model is generated, a second simulation is run with classifier. Simulator selects the lowest possible SF considering prediction of transmission. If a transmission of a node with lowest possible SF is predicted as interfered transmission by classifier, then simulator assigns a higher SF. If no SF is predicted as successful, then simulator again assigns lowest possible SF.

@@

\par In Figure \ref{fig:sf_pdr}, PDR plot of different SF assignment is shown. Randomly generated network topology radius is set to 3000 meters, number of GW is set to 1 and packet generation rate is set to 0.01 packet per second. Increasing SFs increases time on air. This increases the number of collisions thus decreases the PDR of the network. PDR of higher SFs are decreased to almost 0 while number of nodes is increasing. Since network topology radius is quite small almost all SFs can reach to the GW.

@@

\par In Figure \ref{fig:predict_pdr}, PDR plot of machine learning prediction scheme is shown. Randomly generated network topology radius is set to 5000 meters, number of GW is set to 3 and packet generation rate is set to 0.01 packet per second. Prediction model needs nodes location so three GWs are used for prediction SF scheme since three GWs are enough to roughly locate nodes positions by triangulation. Prediction scheme gives better PDR when number of nodes increases since number of interferences also increases. Prediction scheme can only improve network performance when LoRa interference is present. Also, prediction scheme gives better results when nodes are deployed close to the gateway, since nodes have margin to increase their SFs. If a node is far away from the GW, then prediction scheme cannot increase the SF to avoid interference since the SF is already high.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{predict_pdr_r5000_g3_p0.01_s3600}.png}
\caption{PDR of SF prediction scheme.}
\label{fig:predict_pdr}
\end{figure}

@@

\par As for future works, transmit power optimization can be included to proposed machine learning scheme. In this paper, we assume that nodes always use maximum transmit power for uplink transmission, however nodes close to the GW can decrease transmit power to save energy. This will make transmissions to more vulnerable to interference thus requires extra care. Optimizing another node parameter alongside SF can be interesting research. Also, other machine learning methods can be investigated for SF optimization. Reinforcement learning is a good candidate for this network optimization issue.
